---
title: "Figures"
author: "Angela Crabtree"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      root.dir=dirname(rstudioapi::getSourceEditorContext()$path))
```

```{r libraries}
library(RColorBrewer)
library(gplots)
library (ggplot2)
library(cowplot)
library (ggfortify)
library(ggrepel)
library(dplyr)
library(svglite)
library(kableExtra)
```

```{r import_custom_functions, eval=FALSE}
# import custom functions
source("dsrna-survey-functions.R")
```


```{r set_wd}
## set working directory
script_dir=dirname(rstudioapi::getSourceEditorContext()$path)
data_dir=paste0(unlist(strsplit(script_dir, "/scripts"))[1], "/data/")
setwd(data_dir)
## set image output directory
img_dir=paste0(unlist(strsplit(script_dir, "/scripts"))[1], "/img/")
## data files
killer_data_file = paste0(data_dir, "killer_assays.csv")
```


## Results: Part 1

#### **Prevalence of killer yeasts**

This figure contains the killing patterns of four S. cerevisiae strains we've identified as being "canonical": BJH001 (K1 toxin), NCYC1001 (K2 toxin), MS300c (K28 toxin), and DSM 70459 (Klus toxin). 

```{r fig1b}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# only keep canonical killers
killer.df <- killer.df[which(killer.df$strain_name=="BJH001" |    
                               killer.df$strain_name=="NCYC1001" |    
                               killer.df$strain_name=="MS300c" |    
                               killer.df$strain_name=="DSM 70459" ),]
# create matrix to use for heat map
Toxin <- killer.df$toxin
Strain <- killer.df$strain_name
killer.df[,9:20] <- lapply(killer.df[,9:20], as.integer)
killerMatrix <- as.matrix(killer.df[, 9:20])
rownames(killerMatrix) <- Toxin
```

```{r, eval=FALSE}
## save heat map to file
img_file = paste0(img_dir, "Fig1B_canonical_heatmap.svg") # filename
svglite(filename=img_file, 
    width=9, 
    height=5)
# create heat map (ignore error- it still prints to file anyway)
myColors = c('oldlace','steelblue')
canonical_heatmap = heatmap.2( 
  killerMatrix, 
  dendrogram = 'none', Rowv = TRUE, Colv = FALSE,
  trace = 'none', margins = c(8, 6), lwid = c(0.2,5),
  lhei = c(0.2,5), key = FALSE, col = myColors,
  xlab='Lawns', ylab="Killers", colsep = 0:ncol(killerMatrix), 
  rowsep=0:nrow(killerMatrix),
  sepcolor='white', sepwidth=c(0.02, 0.02), 
  main="Toxin Controls")
```

This figure contains the killing patterns of all *S. cerevisiae* strains we screened in the collections which killed at least one lawn that was not Y2046 (because that lawn is hypersusceptible) or ATCC2001 (a lawn we screened but did not include since is not S. cerevisiae). 

```{r fig1d}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove Y2046 as a lawn (Y2046 column) because it is hypersusceptible
killer.df = subset(killer.df, select = -c(Y2046) )
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y1088!=0 |  
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
# create matrix to use for heat map
Toxin <- killer.df$toxin
Strain <- killer.df$strain_name
killer.df$num_strains = 1
killer.df.collapsed = aggregate(num_strains ~ Y1088+DSM70459+CYC1058+Y27106+NCYC1006+NCYC777+NBRC1802+NBRC1815+Y5509+MS300c+BY4741, 
                      data=killer.df, FUN=sum)
killer.df.agg.strain = aggregate(strain_name ~ Y1088+DSM70459+CYC1058+Y27106+NCYC1006+NCYC777+NBRC1802+NBRC1815+Y5509+MS300c+BY4741, 
                      data=killer.df, paste, collapse=",")
killer.df.collapsed$strain_name = killer.df.agg.strain$strain_name
killer.df.collapsed$group = seq(1:nrow(killer.df.collapsed))
K1 = as.numeric(grep('BJH001', killer.df.collapsed$strain_name))
K2 = as.numeric(grep('NCYC1001', killer.df.collapsed$strain_name))
K28 = as.numeric(grep('MS300c', killer.df.collapsed$strain_name))
Klus = as.numeric(grep('DSM 70459', killer.df.collapsed$strain_name))
K21 = as.numeric(grep('T21.4 OS40', killer.df.collapsed$strain_name))
K45 = as.numeric(grep('N-45 OS78', killer.df.collapsed$strain_name))
K62 = as.numeric(grep('Q62.5 OS169', killer.df.collapsed$strain_name))
K74a = as.numeric(grep('Q74.4 OS294', killer.df.collapsed$strain_name))
K74b = as.numeric(grep('Y8.5 OS299', killer.df.collapsed$strain_name))
# rename groups according to previously identified toxin-containing strains
killer.df.collapsed$group[K1]=paste(K1, "(K1)")
killer.df.collapsed$group[K2]=paste(K2, "(K2)")
killer.df.collapsed$group[K28]=paste(K28, "(K28)")
killer.df.collapsed$group[Klus]=paste(Klus, "(Klus)")
killer.df.collapsed$group[K21]=paste(K21, "(K21)")
killer.df.collapsed$group[K45]=paste(K45, "(K45)")
killer.df.collapsed$group[K62]=paste(K62, "(K62)")
killer.df.collapsed$group[K74a]=paste(K74a, "(K74)")
killer.df.collapsed$group[K74b]=paste(K74b, "(K74)")
killer.df.collapsed$group = paste("Group ", killer.df.collapsed$group, 
                                  ", ", killer.df.collapsed$num_strains, " strains", sep="")
# collapse killer strains that killed the exact same strains (consolodates into groups)
nlawns = ncol(killer.df.collapsed)-3
killer.df.collapsed[,1:nlawns] <- lapply(killer.df.collapsed[,1:nlawns], as.integer)
killerMatrix <- as.matrix(killer.df.collapsed[,1:nlawns])
rownames(killerMatrix) <- killer.df.collapsed$group
```

```{r, eval=FALSE}
## save heat map to file
img_file = paste0(img_dir, "Fig1D_collapsed_heatmap.svg") # filename
svglite(filename=img_file, 
    width=6, 
    height=18)
# create heat map (ignore error- it still prints to file anyway)
myColors = c('oldlace','steelblue')
heatmap.2(
  killerMatrix, 
  dendrogram = 'row', Rowv = TRUE, Colv = FALSE,
  trace = 'none', margins = c(8, 10), lwid = c(0.2,5),
  lhei = c(0.2,5), key = FALSE, col = myColors,
  xlab='Lawns', ylab="Killer Groups", colsep = 0:ncol(killerMatrix), 
  rowsep=0:nrow(killerMatrix),
  sepcolor='white', sepwidth=c(0.01, 0.01), 
  main="Collapsed Heatmap")
```

This table does include Y2046 as a killer and only uses strains that we have gel data for from the collections, so we can say whether the strain appears to have dsRNAs. "NGS" strains are not included here. If bands on gels were considered to be narnavirus, they did not count as dsRNA since they are ssRNA and only transiently become dsRNA during replication. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df.stats=killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# Limit query to only those strains that have agarose gel data
killer.df.stats = killer.df.stats[which(killer.df.stats$tv!="NA"),]
# print names of collections included in dataset
print("Names of included collections:")
unique(killer.df.stats$Collection)
# ensure total number of strains in collection subsets adds up correctly
print(paste("Number of qualified strains:", nrow(killer.df.stats))) 
```

```{r tbl1}
#######################################################################
# Calculate numbers of strains in each prevalence category
category = c("No dsRNAs", "dsRNAs", "    Totivirus only", 
             "    Totivirus and satellite(s)", "    Satellites only")
#######################################################################
## KILLERS
killer.df = killer.df.stats
killer.df <- killer.df[which(
  killer.df$Y2046!=0 |
    killer.df$Y1088!=0 |  
    killer.df$DSM70459!=0 |
    killer.df$CYC1058!=0 |  
    killer.df$Y27106!=0 | 
    killer.df$NCYC1006!=0 |
    killer.df$NCYC777!=0 | 
    killer.df$NBRC1802!=0 | 
    killer.df$NBRC1815!=0 |
    killer.df$Y5509!=0 | 
    killer.df$MS300c!=0 |
    killer.df$BY4741!=0),]
# killer, no dsRNAs
kill_ndsrna = killer.df[which(
  killer.df$tv==0 & 
    killer.df$sat==0 &
    killer.df$pv==0),]
# killer, containing any dsRNAs
kill_dsrna = killer.df[which(
  killer.df$tv!=0 | 
    killer.df$sat!=0 |
    killer.df$pv!=0),]
# killer, containing only totivirus
kill_onlytoti = killer.df[which(
  killer.df$tv!=0 & 
    killer.df$sat==0 &
    killer.df$pv==0),]
# killer, containing totivirus & satellite dsRNA
kill_totisat = killer.df[which(
  killer.df$tv!=0 & 
    (killer.df$sat!=0 |
    killer.df$pv!=0)),]
# killer, containing only satellite dsRNA
kill_onlysat = killer.df[which(
  killer.df$tv==0 & 
    (killer.df$sat!=0 |
    killer.df$pv!=0)),]
# combine all killer numbers into list 
killers = as.numeric(c(
  nrow(kill_ndsrna), nrow(kill_dsrna), nrow(kill_onlytoti), nrow(kill_totisat),
  nrow(kill_onlysat)))
#######################################################################
## NON-KILLERS
nonkiller.df = killer.df.stats
nonkiller.df <- nonkiller.df[which(
  nonkiller.df$Y2046==0 &
    nonkiller.df$Y1088==0 &
    nonkiller.df$DSM70459==0 &
    nonkiller.df$CYC1058==0 &
    nonkiller.df$Y27106==0 &
    nonkiller.df$NCYC1006==0 &
    nonkiller.df$NCYC777==0 &
    nonkiller.df$NBRC1802==0 &
    nonkiller.df$NBRC1815==0 &
    nonkiller.df$Y5509==0 &
    nonkiller.df$MS300c==0 &
    nonkiller.df$BY4741==0),]
# non-killer, no dsRNAs
nokill_ndsrna = nonkiller.df[which(
  nonkiller.df$tv==0 & 
    nonkiller.df$sat==0 &
    nonkiller.df$pv==0),]
# killer, containing any dsRNAs
nokill_dsrna = nonkiller.df[which(
  nonkiller.df$tv!=0 | 
    nonkiller.df$sat!=0 |
    nonkiller.df$pv!=0),]
# killer, containing only totivirus
nokill_onlytoti = nonkiller.df[which(
  nonkiller.df$tv!=0 & 
    nonkiller.df$sat==0 &
    nonkiller.df$pv==0),]
# killer, containing totivirus & satellite dsRNA
nokill_totisat = nonkiller.df[which(
  nonkiller.df$tv!=0 & 
    (nonkiller.df$sat!=0 |
    nonkiller.df$pv!=0)),]
# killer, containing only satellite dsRNA
nokill_onlysat = nonkiller.df[which(
  nonkiller.df$tv==0 & 
    (nonkiller.df$sat!=0 |
    nonkiller.df$pv!=0)),]
# combine all killer numbers into list 
nonkillers = as.numeric(c(
  nrow(nokill_ndsrna), nrow(nokill_dsrna), nrow(nokill_onlytoti), 
  nrow(nokill_totisat), nrow(nokill_onlysat)))
#######################################################################
# Create prevalence data frame
prevalence <- data.frame(category, killers, nonkillers)
prevalence$totals = prevalence$killers + prevalence$nonkillers
# print table
prevalence %>% 
  rename(c(` `=category, `Killer yeast`=killers, 
           `Non-killer yeast`=nonkillers, Total=totals)) %>% 
  kbl(caption = "Table 1. The prevalence of dsRNAs in killer and non-killer yeasts in S. cerevisiae") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

## Results: Part 2

#### **Identification of unknown toxins**

The table shows a numerical comparison between the de novo assembled sequences from NGS with those of the canonical reference sequences. 

```{r data_pt2}

```

```{r tbl2}
# Create empty data frame
ngs_summary <- data.frame( 
  strain = character(),    
  toxin = character(),
  ident = character(),
  cov = character(), 
  pcr = character(),
  stringsAsFactors = FALSE)
# Append row for each NGS strain
ngs_summary[1, ] <- list("YJM1077","K1", ">99%", "97%", "-") 
ngs_summary[2, ] <- list("YJM1287","K1", ">99%", "91%", "-") 
ngs_summary[3, ] <- list("YJM1307","K1", ">99%", "90%", "-") 
ngs_summary[4, ] <- list("YJM1077","K1", "98%", "97%", "yes") 
ngs_summary[5, ] <- list("YJM1482","K1", "99%", ">99%", "yes") 
ngs_summary[6, ] <- list("YJM1490","K1", "98%", "97%", "-") 
ngs_summary[7, ] <- list("YJM1621","K1", "99%", "98%", "yes") 
ngs_summary[8, ] <- list("YJM1619","K28", "92%", "97%", "yes") 
ngs_summary[9, ] <- list("YJM145","Klus", ">99%", "94%", "-") 
# print table
ngs_summary %>% 
  rename(c(`S. cerevisiae`=strain, `Toxin type`=toxin, 
           `Identity (nt)`=ident, Coverage=cov, `Rt-PCR confirmed`=pcr)) %>% 
  kbl(caption = "Table 2. Identification of dsRNA elements from NGS assembly") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

The bubble plot compares PCA when excluding or including Y-2046, a hypersuceptible strain. There are 357 strains that kill this lawn and no others (largest bubble in panel A). 

```{r fig3a_data}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y2046!=0 |
                               killer.df$Y1088!=0 |  
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
# create matrix for PCA
Toxin <- killer.df$toxin
Strain <- killer.df$strain_name
Collection <- killer.df$Collection
identifiers <- as.data.frame(cbind(Toxin,Strain,Collection))
killer.df[,9:20] <- lapply(killer.df[,9:20], as.integer)
killerMatrix <- as.matrix(killer.df[,9:20])
rownames(killerMatrix) <- Strain
# calculate PCA
set.seed(123) # set seed so values remain constant each time
killerPCA <- prcomp(killerMatrix)
# create a PCA dataframe for graphing
pcaDF <- as.data.frame(cbind(identifiers, killerPCA$x))
pcaDF$num_strains = 1
pcaDF.agg = aggregate(num_strains ~ PC1+PC2+Toxin, data=pcaDF, FUN=sum)
pcaDF.agg.strain = aggregate(Strain ~ PC1+PC2+Toxin, 
                                 data=pcaDF, paste, collapse=",")
pcaDF.agg$Strain = pcaDF.agg.strain$Strain
pcaDF.agg$group = seq(1:nrow(pcaDF.agg.strain))
# combine unknowns into known toxin bubbles if PC1&PC2 values are identical
pcaDF.agg.combo.toxin = aggregate(Toxin ~ PC1+PC2, 
                                 data=pcaDF.agg, paste, collapse=",")
Toxin = gsub("(\\w+),unknown", "\\1", pcaDF.agg.combo.toxin$Toxin)
pcaDF.agg.combo.numstrain = aggregate(num_strains ~ PC1+PC2, 
                                 data=pcaDF.agg, sum)
pcaDF.agg.combo.strain = aggregate(Strain ~ PC1+PC2, 
                                 data=pcaDF.agg, paste, collapse=",")
# make new dataframe with combos
pcaDF.agg.combo = data.frame(
  PC1=pcaDF.agg.combo.toxin$PC1, 
  PC2=pcaDF.agg.combo.toxin$PC2, 
  Toxin=Toxin, 
  num_strains=pcaDF.agg.combo.numstrain$num_strains, 
  Strain=pcaDF.agg.combo.strain$Strain)
```

```{r fig3a_plot}
## make bubble plot
propVar <- summary(killerPCA)
firstVar <- round(propVar$importance[2, 1] * 100, 2) # contribution of PC1
secondVar <- round(propVar$importance[2, 2] * 100, 2) # contribution of PC2
# load colors
myPalette =  c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", 
               "#FFFF33", "#A65628", "#F781BF", "#b5b3b3")
set.seed(123) # set seed so geom_text_repel label positions remain constant 
# plot
p1 = ggplot(data=pcaDF.agg.combo, aes(x=PC1, y=PC2, size=num_strains, color=Toxin)) +
  geom_point(alpha=0.5) + 
  scale_colour_manual(values=myPalette) +
  #scale_size_continuous(name="# yeast strains", range=c(2,20), 
                        #breaks=c(1,5,30,max(pcaDF.agg.combo$num_strains))) +
  scale_size_identity(name="# yeast strains", trans="sqrt", guide="legend") +
  theme_classic() +
  labs(title="All strains", 
       x=paste('PC1', ' (',
             as.character(firstVar),
             '%', ')', sep = ''), 
       y=paste('PC2', ' (',
               as.character(secondVar),
               '%', ')', sep = '')) +
  geom_text_repel(data = pcaDF.agg.combo[pcaDF.agg.combo$Toxin!="unknown",], 
                  aes(PC1, PC2, label = Toxin), 
                  color="black", size=3.0, 
                  segment.color = 'grey', show.legend = F) + 
  guides(col = FALSE) + 
  scale_x_continuous(expand = c(0.1, 0))
```

```{r fig3b_data}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# remove Y2046 as a lawn (Y2046 column) because it is hypersuceptible
killer.df = subset(killer.df, select = -c(Y2046) )
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y1088!=0 |  
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
# create matrix for PCA
Toxin <- killer.df$toxin
Strain <- killer.df$strain_name
Collection <- killer.df$Collection
identifiers <- as.data.frame(cbind(Toxin,Strain,Collection))
killer.df[,9:19] <- lapply(killer.df[,9:19], as.integer)
killerMatrix <- as.matrix(killer.df[,9:19])
rownames(killerMatrix) <- Strain
# calculate PCA
set.seed(123) # set seed so values remain constant each time
killerPCA <- prcomp(killerMatrix)
# create a PCA dataframe for graphing
pcaDF <- as.data.frame(cbind(identifiers, killerPCA$x))
pcaDF$num_strains = 1
pcaDF.agg = aggregate(num_strains ~ PC1+PC2+Toxin, data=pcaDF, FUN=sum)
pcaDF.agg.strain = aggregate(Strain ~ PC1+PC2+Toxin, 
                                 data=pcaDF, paste, collapse=",")
pcaDF.agg$Strain = pcaDF.agg.strain$Strain
pcaDF.agg$group = seq(1:nrow(pcaDF.agg.strain))
# combine unknowns into known toxin bubbles if PC1&PC2 values are identical
pcaDF.agg.combo.toxin = aggregate(Toxin ~ PC1+PC2, 
                                 data=pcaDF.agg, paste, collapse=",")
Toxin = gsub("(\\w+),unknown", "\\1", pcaDF.agg.combo.toxin$Toxin)
pcaDF.agg.combo.numstrain = aggregate(num_strains ~ PC1+PC2, 
                                 data=pcaDF.agg, sum)
pcaDF.agg.combo.strain = aggregate(Strain ~ PC1+PC2, 
                                 data=pcaDF.agg, paste, collapse=",")
# make new dataframe with combos
pcaDF.agg.combo = data.frame(
  PC1=pcaDF.agg.combo.toxin$PC1, 
  PC2=pcaDF.agg.combo.toxin$PC2, 
  Toxin=Toxin, 
  num_strains=pcaDF.agg.combo.numstrain$num_strains, 
  Strain=pcaDF.agg.combo.strain$Strain)
```

```{r fig3b_plot}
## make bubble plot
propVar <- summary(killerPCA)
firstVar <- round(propVar$importance[2, 1] * 100, 2) # contribution of PC1
secondVar <- round(propVar$importance[2, 2] * 100, 2) # contribution of PC2
# load colors
myPalette =  c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", 
               "#FFFF33", "#A65628", "#F781BF", "#b5b3b3")
set.seed(123) # set seed so geom_text_repel label positions remain constant 
# plot
p2 = ggplot(data=pcaDF.agg.combo, aes(x=PC1, y=PC2, size=num_strains, color=Toxin)) +
  geom_point(alpha=0.5) + 
  scale_colour_manual(values=myPalette) +
  #scale_size_continuous(name="# yeast strains", range=c(2,20), 
                        #breaks=c(1,5,30,max(pcaDF.agg.combo$num_strains))) +
  scale_size_identity(name="# yeast strains", trans="sqrt", guide="legend") +
  theme_classic() +
  labs(title="No Y-2046", 
       x=paste('PC1', ' (',
             as.character(firstVar),
             '%', ')', sep = ''), 
       y=paste('PC2', ' (',
               as.character(secondVar),
               '%', ')', sep = '')) +
  geom_text_repel(data = pcaDF.agg.combo[pcaDF.agg.combo$Toxin!="unknown",], 
                  aes(PC1, PC2, label = Toxin), 
                  color="black", size=3.0, 
                  segment.color = 'grey', show.legend = F) + 
  guides(col = FALSE) + 
  scale_x_continuous(expand = c(0.1, 0))
```

```{r fig3ab_final_fig}
p = plot_grid(p1, p2, labels = c("A", "B"), ncol=2)
## save plot as SVG file
img_file = paste0(img_dir, "Fig3ab_bubble_plots.svg") # filename
ggsave(
  img_file,
  plot = p,
  width = 12, height = 4)
```

## Results: Part 2

#### **Genome-encoded toxins**

```{r genomic_toxin_data}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections 
# (note that we only have KHR/KHS data for FGSC and Liti)
killer.df.stats=killer.df[which(killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"),]
# Limit query to only those strains that have KHR or KHS data
killer.df.stats = killer.df.stats[which(
  killer.df.stats$khr_type!="NA" |
    killer.df.stats$khs_type!="NA"),]
# print names of collections included in dataset
print("Names of included collections:")
unique(killer.df.stats$Collection)
# ensure total number of strains in collection subsets adds up correctly
print(paste("Number of qualified strains:", nrow(killer.df.stats))) 
```

This table summarizes the prevalence of KHR and KHS in relation to killing phenotype. 

```{r tbl3}
## KILLERS
killer.df = killer.df.stats
killer.df <- killer.df[which(
  killer.df$Y2046!=0 |
    killer.df$Y1088!=0 |  
    killer.df$DSM70459!=0 |
    killer.df$CYC1058!=0 |  
    killer.df$Y27106!=0 | 
    killer.df$NCYC1006!=0 |
    killer.df$NCYC777!=0 | 
    killer.df$NBRC1802!=0 | 
    killer.df$NBRC1815!=0 |
    killer.df$Y5509!=0 | 
    killer.df$MS300c!=0 |
    killer.df$BY4741!=0),]
# killer, containing canonical or mutant (not truncated) KHR
kill_khrP = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# killer, containing canonical or mutant (not truncated) KHS
kill_khsP = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# killer, NOT containing canonical or mutant KHR
kill_khrN = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical"),]
# killer, NOT containing canonical or mutant KHS
kill_khsN = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all killer numbers into list 
killers = c("Killer yeast", 
  nrow(kill_khrP), nrow(kill_khsP), nrow(kill_khrN), nrow(kill_khsN))
#######################################################################
## KILLERS of Y-2046 ONLY
killer.df = killer.df.stats
killer.df <- killer.df[which(
  killer.df$Y2046!=0 &
    killer.df$Y1088==0 &
    killer.df$DSM70459==0 &
    killer.df$CYC1058==0 &
    killer.df$Y27106==0 &
    killer.df$NCYC1006==0 &
    killer.df$NCYC777==0 &
    killer.df$NBRC1802==0 &
    killer.df$NBRC1815==0 &
    killer.df$Y5509==0 &
    killer.df$MS300c==0 &
    killer.df$BY4741==0),]
# Y-2046 killer, containing canonical or mutant (not truncated) KHR
kill_khrP_y2046 = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# Y-2046 killer, containing canonical or mutant (not truncated) KHS
kill_khsP_y2046 = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# Y-2046 killer, NOT containing canonical or mutant KHR
kill_khrN_y2046 = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical" ),]
# Y-2046 killer, NOT containing canonical or mutant KHS
kill_khsN_y2046 = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all killer numbers into list 
killers_y2046 = c("Killer yeast (Y-2046 only)", 
  nrow(kill_khrP_y2046), nrow(kill_khsP_y2046), nrow(kill_khrN_y2046), 
  nrow(kill_khsN_y2046))
#######################################################################
## NON-KILLERS
killer.df = killer.df.stats
killer.df <- killer.df[which(
  killer.df$Y2046==0 &
    killer.df$Y1088==0 &
    killer.df$DSM70459==0 &
    killer.df$CYC1058==0 &
    killer.df$Y27106==0 &
    killer.df$NCYC1006==0 &
    killer.df$NCYC777==0 &
    killer.df$NBRC1802==0 &
    killer.df$NBRC1815==0 &
    killer.df$Y5509==0 &
    killer.df$MS300c==0 &
    killer.df$BY4741==0),]
# non-killer, containing canonical or mutant (not truncated) KHR
nokill_khrP = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# non-killer, containing canonical or mutant (not truncated) KHS
nokill_khsP = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# non-killer, NOT containing canonical or mutant KHR
nokill_khrN = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical" ),]
# non-killer, NOT containing canonical or mutant KHS
nokill_khsN = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all non-killer numbers into list 
nonkillers = c("Non-killer yeast", 
  nrow(nokill_khrP), nrow(nokill_khsP), nrow(nokill_khrN), nrow(nokill_khsN))

#######################################################################
# Create prevalence data frame
prevalence <- data.frame( 
  category = character(), 
  khr_pos = numeric(),    
  khs_pos = numeric(),
  khr_neg = numeric(),    
  khs_neg = numeric())

# Append row for each killer category
prevalence[1, ] <- killers
prevalence[2, ] <- killers_y2046
prevalence[3, ] <- nonkillers

# print table
prevalence %>% 
  rename(c(` `=category, 
           `+KHR1`=khr_pos, `+KHS1`=khs_pos, 
           `-KHR1`=khr_neg, `-KHS1`=khs_neg)) %>% 
  kbl(caption = "Table 3. The prevalence of KHR1 and KHS1 in S. cerevisiae") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r tbl4}
# Limit query to only those strains that have agarose gel data
killer.df.stats = killer.df.stats[which(killer.df.stats$tv!="NA"),]
# ensure total number of strains in collection subsets adds up correctly
print(paste("Number of qualified strains:", nrow(killer.df.stats))) 
#######################################################################

## KILLERS WITH SATELLITES
killer.df = killer.df.stats
killer.df <- killer.df[which( # satellite+ only
    killer.df$sat!=0 |
    killer.df$pv!=0),]
killer.df <- killer.df[which( # killers only
  killer.df$Y2046!=0 |
    killer.df$Y1088!=0 |  
    killer.df$DSM70459!=0 |
    killer.df$CYC1058!=0 |  
    killer.df$Y27106!=0 | 
    killer.df$NCYC1006!=0 |
    killer.df$NCYC777!=0 | 
    killer.df$NBRC1802!=0 | 
    killer.df$NBRC1815!=0 |
    killer.df$Y5509!=0 | 
    killer.df$MS300c!=0 |
    killer.df$BY4741!=0),]
# killer, containing canonical or mutant (not truncated) KHR
kill_khrP_satP = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# killer, containing canonical or mutant (not truncated) KHS
kill_khsP_satP = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# killer, NOT containing canonical or mutant KHR
kill_khrN_satP = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical"),]
# killer, NOT containing canonical or mutant KHS
kill_khsN_satP = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all killer numbers into list 
killers_satP = c("+satellite", 
  nrow(kill_khrP_satP), nrow(kill_khsP_satP), nrow(kill_khrN_satP), 
  nrow(kill_khsN_satP))
#######################################################################

## KILLERS NO SATELLITES
killer.df = killer.df.stats
killer.df <- killer.df[which( # satellite neg only
    killer.df$sat==0 &
    killer.df$pv==0),]
killer.df <- killer.df[which( # killers only
  killer.df$Y2046!=0 |
    killer.df$Y1088!=0 |  
    killer.df$DSM70459!=0 |
    killer.df$CYC1058!=0 |  
    killer.df$Y27106!=0 | 
    killer.df$NCYC1006!=0 |
    killer.df$NCYC777!=0 | 
    killer.df$NBRC1802!=0 | 
    killer.df$NBRC1815!=0 |
    killer.df$Y5509!=0 | 
    killer.df$MS300c!=0 |
    killer.df$BY4741!=0),]
# killer, containing canonical or mutant (not truncated) KHR
kill_khrP_satN = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# killer, containing canonical or mutant (not truncated) KHS
kill_khsP_satN = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# killer, NOT containing canonical or mutant KHR
kill_khrN_satN = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical"),]
# killer, NOT containing canonical or mutant KHS
kill_khsN_satN = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all killer numbers into list 
killers_satN = c("-satellite", 
  nrow(kill_khrP_satN), nrow(kill_khsP_satN), nrow(kill_khrN_satN), 
  nrow(kill_khsN_satN))
#######################################################################

## NON-KILLERS WITH SATELLITES
killer.df = killer.df.stats
killer.df <- killer.df[which( # satellite+ only
    killer.df$sat!=0 |
    killer.df$pv!=0),]
killer.df <- killer.df[which(
  killer.df$Y2046==0 &
    killer.df$Y1088==0 &
    killer.df$DSM70459==0 &
    killer.df$CYC1058==0 &
    killer.df$Y27106==0 &
    killer.df$NCYC1006==0 &
    killer.df$NCYC777==0 &
    killer.df$NBRC1802==0 &
    killer.df$NBRC1815==0 &
    killer.df$Y5509==0 &
    killer.df$MS300c==0 &
    killer.df$BY4741==0),]
# non-killer, containing canonical or mutant (not truncated) KHR
nokill_khrP_satP = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# non-killer, containing canonical or mutant (not truncated) KHS
nokill_khsP_satP = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# non-killer, NOT containing canonical or mutant KHR
nokill_khrN_satP = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical" ),]
# non-killer, NOT containing canonical or mutant KHS
nokill_khsN_satP = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all non-killer numbers into list 
nonkillers_satP = c("+satellite", 
  nrow(nokill_khrP_satP), nrow(nokill_khsP_satP), nrow(nokill_khrN_satP),
  nrow(nokill_khsN_satP))
#######################################################################

## NON-KILLERS NO SATELLITES
killer.df = killer.df.stats
killer.df <- killer.df[which( # satellite neg only
    killer.df$sat==0 &
    killer.df$pv==0),]
killer.df <- killer.df[which(
  killer.df$Y2046==0 &
    killer.df$Y1088==0 &
    killer.df$DSM70459==0 &
    killer.df$CYC1058==0 &
    killer.df$Y27106==0 &
    killer.df$NCYC1006==0 &
    killer.df$NCYC777==0 &
    killer.df$NBRC1802==0 &
    killer.df$NBRC1815==0 &
    killer.df$Y5509==0 &
    killer.df$MS300c==0 &
    killer.df$BY4741==0),]
# non-killer, containing canonical or mutant (not truncated) KHR
nokill_khrP_satN = killer.df[which(
  killer.df$khr_type=="mutant" | 
    killer.df$khr_type=="canonical"),]
# non-killer, containing canonical or mutant (not truncated) KHS
nokill_khsP_satN = killer.df[which(
  killer.df$khs_type=="mutant" | 
    killer.df$khs_type=="canonical"),]
# non-killer, NOT containing canonical or mutant KHR
nokill_khrN_satN = killer.df[which(
  killer.df$khr_type!="mutant" & 
    killer.df$khr_type!="canonical" ),]
# non-killer, NOT containing canonical or mutant KHS
nokill_khsN_satN = killer.df[which(
  killer.df$khs_type!="mutant" & 
    killer.df$khs_type!="canonical" ),]
# combine all non-killer numbers into list 
nonkillers_satN = c("-satellite", 
  nrow(nokill_khrP_satN), nrow(nokill_khsP_satN), nrow(nokill_khrN_satN),
  nrow(nokill_khsN_satN))
#######################################################################

# Create prevalence data frame
prevalence <- data.frame( 
  category = character(), 
  khr_pos = numeric(),    
  khs_pos = numeric(),
  khr_neg = numeric(),    
  khs_neg = numeric())

# Append row for each killer category
prevalence[1, ] <- c("Killer yeast", "", "", "", "")
prevalence[2, ] <- killers_satP
prevalence[3, ] <- killers_satN
prevalence[4, ] <- c("Non-killer yeast", "", "", "", "")
prevalence[5, ] <- nonkillers_satP
prevalence[6, ] <- nonkillers_satN

# print table
prevalence %>% 
  rename(c(` `=category, 
           `+KHR1`=khr_pos, `+KHS1`=khs_pos, 
           `-KHR1`=khr_neg, `-KHS1`=khs_neg)) %>% 
  kbl(caption = "Table 4. The prevalence of KHR1 and KHS1 in killer and non-killer S. cerevisiae lacking satellite dsRNAs") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

## Supplement

```{r data_supplement}

```

```{r figS1}

```

```{r figS4}

```

```{r figS5}

```

```{r figS6}

```

## R Session Info

```{r}
sessionInfo()
```

