---
title: "Context Stats for dsRNA Survey Paper"
author: "Angela Crabtree"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      root.dir=dirname(rstudioapi::getSourceEditorContext()$path))
```

```{r libraries}
library(RColorBrewer)
library(gplots)
library (ggplot2)
library(cowplot)
library (ggfortify)
library(ggrepel)
library(dplyr)
library(svglite)
library(kableExtra)
```

```{r import_custom_functions, eval=FALSE}
# import custom functions
source("dsrna-survey-functions.R")
```


```{r set_wd}
## set working directory
script_dir=dirname(rstudioapi::getSourceEditorContext()$path)
main_dir=unlist(strsplit(script_dir, "/figures/scripts"))[1]
## set image output directory
img_dir=paste0(unlist(strsplit(script_dir, "/scripts"))[1], "/img/")
## data files
data_dir=paste0(main_dir, "/killer_assays/data/")
setwd(data_dir)
killer_data_file = paste0(data_dir, "killer_assays.csv")
```

## Total strains in study

These included yeast strains from the Dudley collection, FGSC, Liti1000, and NGS. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
```

```{r}
total_strains = nrow(killer.df)
print(paste0("Total strains: ", total_strains))
```

## Total killers in collections

From the above number, these are any killers from the 4 collections that killed any lawns, excluding C. glabrata lawn ATCC2001. C. glabrata lawn ATCC2001 information has been excluded in all downstream analyses. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y1088!=0 | 
                               killer.df$Y2046!=0 |
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
```

```{r}
n_killers = nrow(killer.df)
print(paste0("Total killer strains: ", n_killers))
print(paste0("Total killer strains, ratio of total: ", n_killers/total_strains))
```

## Total killers in collections, without Y2046

These are killers from the 4 collections excluding the Y2046 lawn since it is the hypersusceptible strain. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove Y2046 as a lawn (Y2046 column) because it is hypersusceptible
killer.df = subset(killer.df, select = -c(Y2046) )
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y1088!=0 |  
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
```

```{r}
n_killers = nrow(killer.df)
print(paste("Total killer strains:", n_killers))
print(paste("Total killer strains, ratio of total:", n_killers/total_strains))
```

## Strains killing at least Y2046

These are strains from the 4 collections which killed the hypersusceptible Y2046 lawn. These may have only killed Y2046 or other lawns. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# only keep strains that killed at least one lawn
killer.df <- killer.df[which(killer.df$Y2046!=0),]
```

```{r}
n_killers = nrow(killer.df)
print(paste0("Total strains killing Y2046: ", n_killers))
print(paste0("Total strains killing Y2046, ratio of total: ", n_killers/total_strains))
```

## Strains only killing Y2046

These are strains from the 4 collections which killed only the hypersusceptible Y2046 lawn. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# only keep strains that killed Y2046
killer.df <- killer.df[which(killer.df$Y2046!=0),]
# only keep strains that didn't kill anything else
killer.df <- killer.df[which(killer.df$Y1088==0 &  
                               killer.df$DSM70459==0 &  
                               killer.df$CYC1058==0 &  
                               killer.df$Y27106==0 &  
                               killer.df$NCYC1006==0 &  
                               killer.df$NCYC777==0 &  
                               killer.df$NBRC1802==0 &  
                               killer.df$NBRC1815==0 &  
                               killer.df$Y5509==0 &  
                               killer.df$MS300c==0 &  
                               killer.df$BY4741==0),]
```

```{r}
n_killers = nrow(killer.df)
print(paste0("Total strains killing only Y2046: ", n_killers))
print(paste0("Total strains killing only Y2046, ratio of total: ", n_killers/total_strains))
```

## Strains containing dsRNAs and/or satellites

These strains contain dsRNAs, and some specifically with satellites. 

```{r}
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# Limit query to only those strains that have agarose gel data
killer.df = killer.df[which(killer.df$tv!="NA"),]
n_gels = nrow(killer.df)

# number of strains containing dsRNA
killer.df.dsrna = killer.df[which(killer.df$tv!=0 | 
                                    killer.df$sat!=0 |
                                    killer.df$pv!=0),]
n_strains_dsrna = nrow(killer.df.dsrna)

# only keep strains that killed at least one lawn
killer.df.killers <- killer.df[which(killer.df$Y1088!=0 | 
                              killer.df$Y2046!=0 |
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
n_killers = nrow(killer.df.killers)
killer.df.killers = killer.df.killers[which(killer.df.killers$tv!=0 | 
                                    killer.df.killers$sat!=0 |
                                    killer.df.killers$pv!=0),]
n_dsrna_killers = nrow(killer.df.killers)
```

```{r}
print(paste("Number of strains tested for dsRNAs (gels):", n_gels))
print(paste("Number of dsRNA+ strains out of those tested:", n_strains_dsrna))
print(paste("Number of dsRNA+ strains out of those tested, ratio:", n_strains_dsrna/n_gels))
print(paste("Number of dsRNA+ out of killer strains:", n_dsrna_killers))
print(paste("Number of dsRNA+ out of killer strains, ratio:", n_dsrna_killers/n_killers))
```

## Strains containing dsRNAs and/or satellites, without Y2046

These strains contain dsRNAs, and some specifically with satellites, excluding hypersusceptible strain Y2046. 

```{r}
# import data
killer.df <- read.csv(killer_data_file, header = TRUE, na.strings = c('*'))
# remove rows where yeast did not grow
killer.df <- killer.df[which(killer.df$Y1088!="no yeast"),]
# remove Y2046 as a lawn (Y2046 column) because it is hypersusceptible
killer.df = subset(killer.df, select = -c(Y2046) )
# remove ATCC2001 as a lawn (ATCC2001 column) because it is not S. cerevisiae
killer.df = subset(killer.df, select = -c(ATCC2001) )
# only include specified collections
killer.df <- killer.df[which(killer.df$Collection=="Dudley" |
                       killer.df$Collection=="FGSC" |
                       killer.df$Collection=="Liti1000"|
                       killer.df$Collection=="NGS"),]
# Limit query to only those strains that have agarose gel data
killer.df = killer.df[which(killer.df$tv!="NA"),]
n_gels = nrow(killer.df)

# number of strains containing dsRNA
killer.df.dsrna = killer.df[which(killer.df$tv!=0 | 
                                    killer.df$sat!=0 |
                                    killer.df$pv!=0),]
n_strains_dsrna = nrow(killer.df.dsrna)

# only keep strains that killed at least one lawn
killer.df.killers <- killer.df[which(killer.df$Y1088!=0 | 
                               killer.df$DSM70459!=0 |
                               killer.df$CYC1058!=0 |  
                               killer.df$Y27106!=0 | 
                               killer.df$NCYC1006!=0 |
                               killer.df$NCYC777!=0 | 
                               killer.df$NBRC1802!=0 | 
                               killer.df$NBRC1815!=0 |
                               killer.df$Y5509!=0 | 
                               killer.df$MS300c!=0 |
                               killer.df$BY4741!=0),]
n_killers = nrow(killer.df.killers)
killer.df.killers = killer.df.killers[which(killer.df.killers$tv!=0 | 
                                    killer.df.killers$sat!=0 |
                                    killer.df.killers$pv!=0),]
n_dsrna_killers = nrow(killer.df.killers)
```

```{r}
print(paste("Number of dsRNA+ out of killer strains:", n_dsrna_killers))
print(paste("Number of dsRNA+ out of killer strains, ratio:", n_dsrna_killers/n_killers))
```

